The five pillars of computational reproducibility

### 2023-05-18 docs/5pillars.Rmd
data sharing

Another challenge is the fact that different parts of the workflow have different compute needs.
For example, genomic sequencing analysis workflows typically involve the mapping of reads to the
reference genome and variant calling which are demanding in memory and compute.
These computationally demanding steps are typically conducted on high end servers and workstations.
After initial processing, the resulting variant files are ~100X smaller than the raw sequence files,
so downstream analysis could be conducted on commodity PCs or laptops.
In those situations,

Later stage

@@

It is not sufficient to simply use these tools, they need to be shared as part of the publication
process.
Conda environments are described by a file called `environment.yml` which acts like a recipe for how to
make the environment.
As they are a recipe, these are small files which are easily shared in the project code repository.
Docker images are larger, and cannot be shared in the code repository.
They are commonly deposited to the DockerHub repository, but availability of images is dependent on the
commercial viability of Docker Inc, so it isn't guaranteed to survive the 10-year challenge.
A better option for bioinformatics workflows would be to use the BioContainers registry
@Da_Veiga_Leprevost2017-yw, as it is community driven and supported by consortia such as *Elixir* and
*Global Alliance for Genomics and Health* that will underpin its longevity.
Still, these are not considered permanent archival services, so it is advised to deposit the image
used in anarticle to a suitable persitent long-term archive (discussed below).
The recipe used to build the Docker image, called the Dockerfile, is sufficiently small and can be added
to the project repository.

@@

@@

https://www.minot.bio/
https://us02web.zoom.us/rec/play/zZsJzFME-9Uu5S5BKwX-9DRXQfHq1wnZjH4Mrgy2v1l3e2iE9BkA_Bwj8ZDcoz19BADqmZ6KU3hLA-oP.RKy6CJCCZD_iRjV2?canPlayFromShare=true&componentName=rec-play&continueMode=true&from=share_recording_detail&originRequestUrl=https%3A%2F%2Fus02web.zoom.us%2Frec%2Fshare%2FRbeIXIFsAd_oFNttD7zBeC-W8vswX-rq3r4XoUmlp2V2RFJLTd_wut9qhfuzrESF.wM6AyVmmcQks_cdy

>>>>

@@

It is not sufficient to simply use these tools, they need to be shared as part of the publication
process.
Conda environments are described by a file called `environment.yml` which acts like a recipe for how to
make the environment.
As they are a recipe, these are small files which are easily shared in the project code repository.
Docker images are larger, and cannot be shared in the code repository.
They are commonly deposited to the DockerHub repository, but availability of images is dependent on the
commercial viability of Docker Inc, so it isn't guaranteed to survive the 10-year challenge.
A better option for bioinformatics workflows would be to use the BioContainers registry
@Da_Veiga_Leprevost2017-yw, as it is community driven and supported by consortia such as *Elixir* and
*Global Alliance for Genomics and Health* that will underpin its longevity.
Still, these are not considered permanent archival services, so it is advised to deposit the image
used in an article to a suitable persitent long-term archive (discussed below).
The recipe used to build the Docker image, called the Dockerfile, is sufficiently small and can be added
to the project repository.

@@

FAIR data sharing is one of the key features of open science, which characterises itself as science
that is collaborative, transparent, accessible and inclusive [@noauthor_2023-uq].

The Universal Declaration of Human Rights states [@United_Nations1948-ua]:

>Everyone has the right freely to participate in the cultural life of the
>community, to enjoy the arts and to share in scientific advancement and
>its benefits.

In this context, access to research data may be interpreted as a benefit of scientific advancement.
In particular, when large data sets are generated with public funds, the public might expect that the
benefts of that study, like the publications and underlying data are released.

Data sharing has dual purposes, firstly this data may be useful in answering other research questions,
and secondly to facilitate reproduction and auditing [@Wilkinson2016-qq].
Data reuse prevents redundant research and facilitates new research ideas that were impossible without
it.
In terms of research rigour, inspecting raw data can uncover research integrity problems [@Miyakawa2020-yj],
in addition to validation of scientific findings by independent computational reproduction [@Ioannidis2009-hv].
In light of this, a research article without supporting data and code is much like a press release or
advertisement and the claims made cannot be verified as true or not.

Although it is common to see journal articles with "Data available upon reasonable request", this is less
than ideal.
I, and many fellow researchers have had the displeasure of attempting to contact corresponding
authors of these studies and learn that they have moved on to another institute, retired, passed away,
or the data has been lost or otherwise unavailable (if they receive a response at all).
In the rare cases that the data has been retained, arranging access can be tedious as such datasets can
run into hundreds of gigabytes which cannot be easily transferred across the internet.
If many requests are made for a dataset, this can become time intensive for the sharers which is
wasteful considering they could have shared the data from the start and reached a larger audience.

FAIR principles: @Wilkinson2016-qq

Long term data archiving (and sharing) not only faciliitates this,

Using secondary data is

Moreover there is

Data available on request?

NIH data sharing requirements.

@@

https://www.minot.bio/
